---
title: "__ placeholder __"
author: "Alex Stern"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Likelihood models under evolutionary asymmetry only

So far we have looked primarily at the power asymmetry model, but we have yet to discuss evolutionary asymmetry.

Under this model, we can calculate the proportional probability of observing a given frequency of an allele based on its selection coefficient. Note that this probability should not be interpreted directly, as it is only a proportional probability for various frequencies among the risk-increasing alleles.

```{r}
## Proportional probability of observing an increasing allele at frequency p, given a selection coefficient
p_x <- function(p, gamma) {
  
  # p = risk-increasing allele frequency
  # gamma = population scale selection coefficient
  
  numerator <- exp(-gamma * p)
  
  denominator <- p * (1 - p)
  
  return(numerator / denominator)
}
```

We can apply Bayes' theorem similarly to how we applied it to the power asymmetry model to calculate the probability of a given sign in the evolutionary model. Most of the terms end up cancelling, leaving us with with a simple equation.

```{r}
# Function to calculate the probability that the minor allele is trait increasing, given its frequency and the selection coefficient
prob_sign_evo <- function(x, gamma) {
  
  # x = minor allele frequency
  # gamma = population scale selection coefficient
  
  result <- 1 / (1 + exp(gamma * (2*x - 1)))
  return(result)
}
```

Using these probabilities, we can simulate data for a trait with a given selection coefficient and minor allele frequency at each site.

```{r}
sim_data_evo <- function(L, vec_x, gamma) {
  
  # L = number of sites
  # x = minor allele frequency
  # gamma = population scale selection coefficient
  
  prob_sign <- rep(0,L)
  for (i in 1:L) {
    prob_sign[i] <- prob_sign_evo(vec_x[i], gamma)
  }
  
  # Draw the sign based on the probability above
  vec_obs_b <- rbinom(L, 1, prob_sign)
  # Convert each element to 1 or -1 depending on whether it increases or decreases the trait
  vec_obs_b <- (vec_obs_b - 0.5) * 2
  
  return(vec_obs_b)
}
```

